Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-26 article.tex
appendix : directivity

The heuristic is based on the following observation. The rise of fluorescence of a neuron indicates its activation.

>>>>

The heuristic is based on the following observation. The rise of fluorescence
of a neuron indicates its activation. If another neuron is activated after a
slight delay, this could have been caused by the activation of the first
neuron and therefore highlight a (direct) directed relationship between the
two neurons. With that in mind, the heuristic principle is to count the number
of such events occurred between neurons $i$ and $j$ for consecutive time
steps. It can be done by counting the number of times that a transition from
neuron $i$ to neuron $j$ for two consecutive time steps, or mathematically we
compute for each pair $(i,j)$

\begin{align}
s_{ij} = \sum_{t=1}^{T - 1} \mathbb{1}(x_j^{t+1} - x_i^t \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic.

Two issues may be spotted. First, this heuristic is not able to distinguish a
directed relationships if both neurons are activated in the same time step and
may wrongly characterize indirect relationships because of the time
resolution. Second, it may be necessary to take into account $s_{ji}$ to
estimate the quality of the pairwise coefficient $s_{i,j}$. Indeed, if both
coefficients are similar ($s_{ij} \sim s_{ji}$) the directivity can not be
inferred since the order of fluorescent rising is not well defined. On the
contrary, different coefficients $s_{ij}$ and $s_{ji}$ mean that an order of
fluorescent rising is well established and observable. There we choose to
compute a differential pairwise coefficient $z_{ij}$ defined by the following
equation

\begin{align}
z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_j^{t+1} - x_i^t \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}(x_i^{t+1} - x_j^t \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

At the light of the poor results of this heuristic taken on its own summarized
on Table \ref{tab:} as \textit{directivity}, we use it to make asymmetric the
degree of association matrix using partial correlation statistics. We expect
that when the heuristic is highly confident (i.e., $z_{i,j} \gg 0 \gg
z_{j,i}$) the degree of association   of edge $(i,j)$ is increased while the
degree of association corresponding to the edge $(j,i)$ is reduced. Even if
this adjustment is subtle, it might have a slight positive effect on the ROC
AUC.
